---
title: "202207_saed_covid_asd"
author: "tai"
date: '2022-07-05'
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())

library(tidyverse)
library(rms)
library(MASS)
library(lme4)
library(ggpubr)
library(ggsci)
```

# init
- read data
```{r}
df <- readxl::read_xls("base_P_C (to RT)_r2.xls")
df_cbcl <- readxl::read_xls(
  "CBCL_rev.xls"
)
source("get_model_stat_202205.r")
```

- handle CBCL
```{r}
df_cbcl_temp <- df_cbcl %>%
  dplyr::select(id, sex, CBCL_4_5_F:CBCL_6_11_M_ex) %>%
  pivot_longer(
    cols = CBCL_4_5_F:CBCL_6_11_M_ex, names_to = "colname",values_to = "value" 
  )
df_cbcl_temp <- df_cbcl_temp %>%
  mutate(
    agegroup = gsub("[a-zA-Z_]*","",gsub("CBCL_","",x = colname)),
    type = gsub("_bl","",gsub("[0-9A-Z]_*","",x = colname))
  ) 

# Chinses normal model in the literature; ref: Su1996
Su1996_mod <- data.frame(
  type = c(
    rep(c("withd", "deps", "soma", "agres",
    "in", "ex"), 8)
  ),
  sex = c(
    rep(0,12), rep(1,12),
    rep(0,12), rep(1,12)
  ),
  agegroup = rep(
    rep(c(rep("45",6), rep("611",6)),4)
  ),
  cbcl_mean = c(
    1.02, 3.45, 1.00, 6.34, 7.31, 8.21, #M, 4-5
    1.08, 1.68, 1.01, 4.55, 6.06, 8.74, #M, 6-11
    1.68, 1.60, 3.59, 4.97, 7.82, 8.17, #F, 4-5
    1.85, 2.46, 1.69, 4.86, 5.26, 8.20 #F, 6-11
  ),
  cbcl_sd = c(
    1.31, 3.16, 1.23, 4.07, 5.76, 5.49, #M, 4-5
    1.19, 2.12, 1.63, 4.10, 5.74, 6.80,  #M, 6-11
    1.79, 2.26, 2.74, 4.02, 6.46, 5.74, #F, 4-5
    1.95, 2.57, 2.25, 4.19, 5.23, 6.55 #F, 6-11
  ),
  stringsAsFactors = F
)

df_cbcl_temp <- df_cbcl_temp %>%
  left_join(Su1996_mod, by = c("type", "sex", "agegroup"))
df_cbcl_temp <- df_cbcl_temp %>%
  distinct(id, sex, colname, .keep_all = T) %>%
  filter(!is.na(cbcl_mean), !is.na(value))
df_cbcl_temp <- df_cbcl_temp %>%
  mutate(z_hnmod = (value-cbcl_mean)/cbcl_sd)
df_cbcl_temp <- df_cbcl_temp %>%
  mutate(r3 = stringr::str_sub(colname, -3,-1),
         bl = case_when(r3=="_bl"~"_bl", TRUE~""),
         type_cat = paste0("Z_HN_",type, bl))

df_cbcl_syn <- df_cbcl_temp %>%
  dplyr::select(id,type_cat,z_hnmod) %>%
  pivot_wider(id_cols = "id", names_from = type_cat, values_from = z_hnmod)

df <- left_join(df, df_cbcl_syn, by = "id")

```

- handle covid-related stress
```{r}
df <- df %>%
  mutate(Stress_total= Stress_1+Stress_2+Stress_3+Stress_4+Stress_5+Stress_6+Stress_7)
df$Stress_total <- as.numeric(df$Stress_total)

df <- df %>%
  mutate(Lock_infection_3cat = case_when(
    infect6==1 & Lock_2 == 0 ~ "0_open_lowrisk",
    infect6==1 & Lock_2 == 1 ~ "1_lock_lowrisk",
    infect6==0 & Lock_2 == 1 ~ "2_lock_highrisk"
  ),Lock_infection_3cat = factor(Lock_infection_3cat)
  )

```

- exposure: lockdown
```{r}
df <- df %>%
  mutate(Lock_days_cat = case_when(
    Lock_2==0 ~ "0_No_Lock",
    Lock_2==1 & Day_lock_2 <35 ~ "1_<35days",
    Lock_2==1 & Day_lock_2 >=35 ~ "2_>=35days"
  ),
  Lock_days_cat = factor(Lock_days_cat),
  Lock_2 = factor(Lock_2)
  )
```

- define variables
```{r}
EXPOSURE_VAR <- c(
  "Lock_2", # 2-cat
  "Lock_days_cat" # 3-cat
)

FACTOR_VAR <- c(
  "Lock_2", "Lock_days_cat",
  "school_influ",
  "behav_influ3", 
  "med_influ2",
  "other_influ2",
  "sex", 
  "Educa_2",
  "Mari_2",
  "Income_2",
  "infect6",
  colnames(df)[startsWith(colnames(df),"Stress_")],
  "Fami_cohes_cat", 
  "Fami_adapt_cat"
)

COVARS <- c(
  "sex", "age",
  "Educa_2",
  "Income_2"
)

COVARS_FIT <- c( # no transformation
  "sex", "age",
  "Educa_2",
  "Income_2"
)

OUTCOME_VAR_FOLLOWUP <- c(
  "srs_aware", "srs_cog", "srs_com", "srs_motiv", "srs_mann",
  "srs_tot",
  "sdq_emo", "sdq_beh", "sdq_hyper", "sdq_peer", "sdq_tot",
  "sdq_prosocial", 
  "sleep_habit", "sleep_anx", "sleep_inreg" , "sleep_osa", 
  "sleep_disturb", "sleep_daysleep",   "sleep_awake", 
  "sleep_latency", "sleep_tot",
  "Z_HN_soma",     "Z_HN_deps",     "Z_HN_withd",  
  "Z_HN_agres",    "Z_HN_in",       "Z_HN_ex"
)

OUTCOME_VAR_BASELINE <- c(
  paste0(OUTCOME_VAR_FOLLOWUP, "_bl")
)

OUTCOME_VAR_2 <- c(
  "cgi_i", "cgi_s"
)

MEDIATION_VAR1 <- c(
  colnames(df)[startsWith(colnames(df),"Stress_")], 
  "Stress_total",
  "school_3cat","behav_3cat",
  "med_3cat",
  "infect6"
)

MEDIATION_VAR2 <- c(
  "Sleep_TS", 
  paste0("Sleep_",LETTERS[1:7])
  
)

MODIFIER_VAR <- c(
  "Fami_cohes", "Fami_adapt",
  "Emo_A",
  "Emo_B",
  "Resil_TS",
  paste0("BFI_",LETTERS[1:5]),
  "infect6"
)

CONTINUOUS_VARS <- setdiff(union(EXPOSURE_VAR, MODIFIER_VAR), FACTOR_VAR)

```

```{r}
to_factor <- function(data, var){
  data[[var]][is.na(data[[var]])] <- 99
  data[[var]] <- as.factor(data[[var]])
  return(data)
}

for (i in FACTOR_VAR){
  df <- to_factor(df, i)
}

df$age <- df$age_month/12
df_dataframe <- data.frame(df)

df <- df %>%
  mutate(
    school_baseline = case_when(
      school34==0 ~ 1,
      TRUE ~ 0
    ),
    behav_baseline = case_when(
      behav_influ1==1 ~ 1,
      TRUE ~ 0
    ),
    med_baseline = case_when(
      med_influ1==1 ~ 1,
      TRUE ~ 0
    ),
    school_cov = case_when(
      school_baseline==1 & school_influ==1 ~ 1,
      TRUE ~ 0
    ),
    behav_cov = case_when(
      behav_baseline==1 & school_influ==1 ~ 1,
      TRUE ~ 0
    ),
    med_cov = case_when(
      school_cov==1 & school_influ==1 ~ 1,
      TRUE ~ 0
    ),
    school_3cat = case_when(
      school_baseline==0 ~ "0_not_attending",
      school_baseline==1 & school_cov==0 ~ "1_uninfluenced",
      school_baseline==1 & school_cov==1 ~ "2_influenced"
    ),
    behav_3cat = case_when(
      behav_baseline==0 ~ "0_not_attending",
      behav_baseline==1 & behav_cov==0 ~ "1_uninfluenced",
      behav_baseline==1 & behav_cov==1 ~ "2_influenced"
    ),
    med_3cat = case_when(
      med_baseline==0 ~ "0_not_attending",
      med_baseline==1 & med_cov==0 ~ "1_uninfluenced",
      med_baseline==1 & med_cov==1 ~ "2_influenced"
    ), school_3cat= factor(school_3cat), behav_3cat=factor(behav_3cat), med_3cat=factor(med_3cat)
  )

dd <- datadist(df)
options(datadist = dd)
```
```{r}
df_mixed_1 <- df %>%
  dplyr::select(id, COVARS, EXPOSURE_VAR, MEDIATION_VAR1, MEDIATION_VAR2, 
         MODIFIER_VAR, OUTCOME_VAR_BASELINE)
colnames(df_mixed_1) <- gsub("_bl","",colnames(df_mixed_1))

df_mixed_2 <- df %>%
  dplyr::select(id, COVARS, EXPOSURE_VAR, MEDIATION_VAR1, MEDIATION_VAR2, 
         MODIFIER_VAR, OUTCOME_VAR_FOLLOWUP)

df_mixed <- bind_rows(df_mixed_1, df_mixed_2,.id = "timepoint")
df_mixed$timepoint <- factor(as.numeric(df_mixed$timepoint)-1)

```

# analysis
- table 1
```{r}
table1::table1(as.formula(paste0(
  "~",paste0(c(COVARS, 
               "infect6", "school_3cat","behav_3cat", "med_3cat",
               paste0("BFI_",LETTERS[1:5])),collapse = "+"),"|",
  "Lock_days_cat"
)),
df)
```

- CGI: figure 1a
```{r}
ggplot(df)+
  geom_bar(aes(Lock_days_cat, fill = factor(cgi_i)), position = position_fill(reverse = T))+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 45, hjust = .8))+
 xlab("")+
  ylab("Proportion")+
 ggsci::scale_fill_aaas(name = "CGI_I")+
  scale_x_discrete(labels = c("No lockdown", "Short-term lockdown", "Long-term lockdown"))
```

```{r}
print(
  df %>%
    group_by(Lock_days_cat) %>%
    summarise(mean_cgi_i = mean(cgi_i),
              sd_cgi_i = sd(cgi_i))
)
```

- CGI-I ~ Lockdown (2-cat)
```{r}
f_cgi_i_lockdays_crude <- lm(as.formula(
        paste0(
          "cgi_i", "~",
          "Lock_2"
          )
        ),
      df)
print(summary(f_cgi_i_lockdays_crude))

f_cgi_i_lockdays <- lm(as.formula(
        paste0(
          "cgi_i", "~",
          "Lock_2"," +",
          paste(COVARS_FIT, collapse = "+")
          )
        ),
      df)
    
print(summary(f_cgi_i_lockdays))

```

- CGI-I ~ Lockdown (3-cat)
```{r}
f_cgi_i_lockdays_crude <- lm(as.formula(
        paste0(
          "cgi_i", "~",
          "Lock_days_cat"
          )
        ),
      df)
print(summary(f_cgi_i_lockdays_crude))

f_cgi_i_lockdays <- lm(as.formula(
        paste0(
          "cgi_i", "~",
          "Lock_days_cat"," +",
          paste(COVARS_FIT, collapse = "+")
          )
        ),
      df)
    
print(summary(f_cgi_i_lockdays))

```

- X ~ lockdown (3-cat); mixed model
```{r}
get_tab_beforeafter <- function(df, set_outcome){
  
  main_mixed_eff_rst <- data.frame(
  )
  
  sub_rst <- data.frame(
    outcome = character(),
    stringsAsFactors = F
  )
  
  for (i in "Lock_days_cat"){
    
    for (j in set_outcome){
      df_temp <- df %>%
        filter(!is.na(.data[[j]]), !is.na(.data[[paste0(j,"_bl")]]))
      
      df_mixed_1_temp <- df_temp %>%
        dplyr::select(id, COVARS, EXPOSURE_VAR, MEDIATION_VAR1, MEDIATION_VAR2, 
                      MODIFIER_VAR, OUTCOME_VAR_BASELINE)
      colnames(df_mixed_1_temp) <- gsub("_bl","",colnames(df_mixed_1_temp))
  
      df_mixed_2_temp <- df_temp %>%
        dplyr::select(id, COVARS, EXPOSURE_VAR, MEDIATION_VAR1, MEDIATION_VAR2, 
                      MODIFIER_VAR, OUTCOME_VAR_FOLLOWUP)
      
      df_mixed_temp <- bind_rows(df_mixed_1_temp, df_mixed_2_temp,.id = "timepoint")
      df_mixed_temp$timepoint <- factor(as.numeric(df_mixed_temp$timepoint)-1)
      
      temp_f <- lmer(as.formula(
          paste0(
            j, "~",
            i," * timepoint+",
            paste(COVARS_FIT, collapse = "+"),"+",
            "(1|id)"
            )
          ),
        df_mixed_temp)
      
      temp_s <- summary(temp_f)
      
      temp_varno <- c(nrow(temp_s$coefficients)-1, nrow(temp_s$coefficients))
      temp_pval <- pnorm(abs(temp_s$coefficient[temp_varno,"t value"]),lower.tail = FALSE)*2
      
      temp_sum <- df_temp %>%
          group_by(.data[[i]]) %>%
          summarise(m=mean(.data[[paste0(j,"_bl")]]),
                    sd=sd(.data[[paste0(j,"_bl")]]),
                    m2=mean(.data[[j]]),
                    sd2=sd(.data[[j]]))
      
      temp_arr <- c()
      for (k in 1:(length(temp_varno)+1)){
        temp_arr <- c(temp_arr, unlist(temp_sum[k,]))
      }
      
      for (k in 1:length(temp_varno)){
        temp_arr <- c(temp_arr, temp_s$coefficients[temp_varno[k],], "P"=as.numeric(temp_pval[k]))
      }
      
      main_mixed_eff_rst <- bind_rows(
        main_mixed_eff_rst, temp_arr
      )
      
      sub_rst <- bind_rows(
        sub_rst, c(outcome = j)
      )
      
    }
  }
  
  main_mixed_eff_rst <- bind_cols(
    sub_rst, main_mixed_eff_rst
  )
  
  temp_fdr <- p.adjust(c(main_mixed_eff_rst$P...20, main_mixed_eff_rst$P...24), method = "fdr")
  main_mixed_eff_rst$fdr1 <- temp_fdr[1:(length(temp_fdr)/2)]
  main_mixed_eff_rst$fdr2 <- temp_fdr[(length(temp_fdr)/2+1):length(temp_fdr)]
  
  return(main_mixed_eff_rst)

}
```

- SRS ~ Lockdown (3-cat)
```{r}
print(get_tab_beforeafter(df, set_outcome = OUTCOME_VAR_FOLLOWUP[startsWith(OUTCOME_VAR_FOLLOWUP, "srs")]))

```

- figure 1b
```{r}
ggpaired(rename(df, "Baseline" = srs_cog_bl,
                "Follow-up"=srs_cog
                ), 
         "Baseline", "Follow-up", 
         facet.by = "Lock_days_cat", 
         panel.labs = list(Lock_days_cat=c("No lockdown","Short-term lockdown","Long-term lockdown")),
         xlab = "Social Cognition (SRS)",
         ylab = "Score"
         )
```


- SDQ ~ Lockdown (3-cat)
```{r}
print("sdq")
print(get_tab_beforeafter(df, set_outcome = OUTCOME_VAR_FOLLOWUP[startsWith(OUTCOME_VAR_FOLLOWUP, "sdq")]))
```

```{r}
print("CBCL")
print(get_tab_beforeafter(df, set_outcome = OUTCOME_VAR_FOLLOWUP[startsWith(OUTCOME_VAR_FOLLOWUP, "Z_HN")]))
```

```{r}
print("CHRQ")
print(get_tab_beforeafter(df, set_outcome = OUTCOME_VAR_FOLLOWUP[startsWith(OUTCOME_VAR_FOLLOWUP, "sleep")]))
```

- additional adjustment: school/behav/med influence
```{r}
for (i in c("school_3cat","behav_3cat", "med_3cat", "infect6")){
  print(i)
  f_cgi_i_inf1 <- lm(as.formula(
    paste0(
        "cgi_i", "~",
        "Lock_days_cat"," +",
        paste(c(COVARS_FIT, i), collapse = "+")
    )
),
df)

sum_temp <- summary(f_cgi_i_inf1)
print(sum_temp)
}

```

- effect modification
```{r}
centralize <- function(x,na.rm=TRUE) {
    return(x-mean(x,na.rm=na.rm))
  }

get_tab_modification <- function(df, modifier, outcome = "cgi_i"){
  df_centralize <- df %>%
  filter(Lock_days_cat %in% c(
    levels(df$Lock_days_cat)[1],
    levels(df$Lock_days_cat)[3]
  )) %>%
  mutate_at(
    vars(CONTINUOUS_VARS),
    centralize
  )
  mod_eff_rst <- data.frame(
)

  sub_rst <- data.frame(
    stringsAsFactors = F
  )
  
  for (j in modifier){
  
    temp_f <- lm(
      as.formula(
        paste0(
          "cgi_i ~ Lock_days_cat*",
          j,  "+",
          paste(COVARS, collapse = "+")
          )
        ),
      df_centralize
    )
  
    temp_s <- summary(temp_f)
  
    temp_varno <- nrow(temp_s$coefficients)
  
    temp_arr <- c()
  
    for (k in 1:length(temp_varno)){
      temp_arr <- c(temp_arr, temp_s$coefficients[temp_varno[k],])
    }
  
    mod_eff_rst <- bind_rows(
      mod_eff_rst, temp_arr
    )
  
    sub_rst <- bind_rows(
      sub_rst, c(mod = j)
    )
  
  }
  
  mod_eff_rst <- bind_cols(
    sub_rst, mod_eff_rst
  )
  temp_fdr <- p.adjust(mod_eff_rst$`Pr(>|t|)`, method = "fdr")
  mod_eff_rst$fdr <- temp_fdr
  
  return(mod_eff_rst)
  
}

```

- modifier: parental bfi
```{r}
print(get_tab_modification(df, modifier = MODIFIER_VAR[startsWith(MODIFIER_VAR, "BFI")]))
```

- figure 2
```{r}
fmt_pval <- function(p){
  temp <- c()
  for (i in p){
    temp_p <- ifelse(is.character(i),i,
           ifelse(i<0.001, "<0.001", paste0("=",sprintf("%0.3f", i))
                  )
           )
    temp <- c(temp, temp_p)
  }
  return(temp)
}

LOCK_LABELS <- list(
  "0_No_Lock" = "No lockdown",
  "1_<35days" = "Lockdown, <35d",
  "2_>=35days" = "Lockdown, ≥35d"
)


for (i in c(4)){
  
  temp_mod <- paste0("BFI_", LETTERS[1:5])[i]
  temp_exp <- "Lock_days_cat"
  temp_out <- "cgi_i"
  
  temp_text <- ""
  
  lev_exp <- levels(df[[temp_exp]])
  
  temp_syn <- data.frame(stringsAsFactors = F)
  
  temp_nonlinear <- lm(
      as.formula(
        paste0(
          temp_out, "~",
          temp_mod, "*",temp_exp,"+",
          paste(COVARS, collapse = "+")
          )
        ),
      df
    )
  
  temp_sumf <- summary(temp_nonlinear)
  temp_num <- nrow(temp_sumf$coefficients)
  nonlinear_pval <- temp_sumf$coefficients[c(temp_num-1, temp_num),4]
  nonlinear_pval <- fmt_pval(nonlinear_pval)
  
  for (k in 1:length(lev_exp)){
    
    temp_f <- Glm(
      as.formula(
        paste0(
          temp_out, "~",
          temp_mod, "+",
          paste(COVARS, collapse = "+")
          )
        ),family = "gaussian",
      df, subset = df[[temp_exp]]==lev_exp[k]
    )
    
    temp_s <- Predict(temp_f)
    temp_s_mod <- temp_s[startsWith(rownames(temp_s), temp_mod), ]
    temp_s_mod$mod <- lev_exp[k]
    
    temp_syn <- bind_rows(
      temp_syn,
      temp_s_mod, 
    )
    
    temp_sumf <- get_model_stats(temp_f)
    temp_pval <- temp_sumf$coefs$`Pr(>|Z|)`[2]
    temp_pval <- fmt_pval(temp_pval)
    
    temp_betastat <- temp_f$coefficients[2]
    
    temp_text <- paste0(
      temp_text,
      LOCK_LABELS[lev_exp[k]][[1]],", β=", sprintf("%0.3f",temp_betastat),", P", 
      temp_pval,
      ifelse(k==1,"", paste0(", P interaction", 
                             nonlinear_pval[k-1]
                             )
             ),
      "\n"
    )
    
  }
  
  temp_syn$mod <- as.factor(temp_syn$mod)
  
  # png(filename = paste0(
  #   temp_mod,"_",temp_out,".png"
  # ))
  
  print(
      ggplot()+
    geom_point(aes(x = df[[temp_mod]],
                   y = df[[temp_out]],
                   color = df[[temp_exp]]),
               data = df)+
    geom_line(
      aes(x=temp_syn[[temp_mod]],
          y=temp_syn$yhat,
          color = temp_syn$mod)
    )+
    geom_ribbon(
      aes(x=temp_syn[[temp_mod]],
          ymin = temp_syn$lower,
          ymax=temp_syn$upper,
          fill = temp_syn$mod),
          alpha = .1
    )+
    geom_text(
      aes(x = -Inf, y = Inf),
      hjust = 0, vjust = 1,
      label=temp_text
      )+
    xlab(temp_mod)+
    ylab(temp_out)+
    theme_minimal()+
    scale_color_discrete(name = "Lockdown", labels = LOCK_LABELS)+
    scale_fill_discrete(name = "Lockdown", labels = LOCK_LABELS)+
      ylab("CGI improvement")+
      xlab("BFI agreeableness")
    )
  
  # dev.off()
}
```

- modifier: CD-RISK-10
```{r}
print(get_tab_modification(df, modifier = MODIFIER_VAR[startsWith(MODIFIER_VAR, "Resil")]))
```

- modifier: CD-RISK-10
```{r}
print(get_tab_modification(df, modifier = MODIFIER_VAR[startsWith(MODIFIER_VAR, "Fami")]))
```

- modifier: ERQ
```{r}
print(get_tab_modification(df, modifier = MODIFIER_VAR[startsWith(MODIFIER_VAR, "Emo")]))
```

- intercorrelation: bfi and others
```{r}
for (i in c("Resil_TS", "Fami_adapt", "Fami_cohes", "Emo_A", "Emo_B")){
  temp_f <- lm(
    as.formula(
      paste0(
        i,"~BFI_D+",
        paste(c("Educa_2","Income_2"), collapse = "+")
        )
      ),
    df
  )
  temp_s <- summary(temp_f)
  
  print(i)
  print(temp_s$coefficients[2,])
}

for (i in c("Stress_total", "log_Anti_T", "Deps_TS")){
  temp_f <- lm(
    as.formula(
      paste0(
        i,"~BFI_D+",
        paste(c("Educa_2","Income_2"), collapse = "+")
        )
      ),
    filter(df, Lock_days_cat=="2_>=35days")
  )
  temp_s <- summary(temp_f)
  
  print(i)
  print(temp_s$coefficients[2,])
}
```
- additional adjustment for family-related variables
```{r}
for (i in c("Resil_TS", "Fami_adapt", "Fami_cohes", "Emo_A", "Emo_B",
            "Stress_total")){
  temp_f <- lm(
    as.formula(
      paste0(
        "cgi_i~Lock_days_cat*BFI_D+",
        i, "+",
        paste(COVARS, collapse = "+")
        )
      ),
    df
  )
  temp_s <- summary(temp_f)
  temp_pval <- temp_s$coefficients[nrow(temp_s$coefficients),4]
  print(temp_s$coefficients[nrow(temp_s$coefficients),])
}
```


